\documentclass[preprint]{aastex}

% packages for figures
\usepackage{graphicx}
% packages for symbols
\usepackage{latexsym,amssymb}
% AMS-LaTeX package for e.g. subequations
\usepackage{amsmath}

%=====================================================================
% FRONT MATTER
%=====================================================================

\slugcomment{Draft \today}

%=====================================================================
% BEGIN DOCUMENT
%=====================================================================

\newcommand{\beq}{\begin{equation}}
\newcommand{\eeq}{\end{equation}}
\newcommand{\shera}{{\sc shera}}
\newcommand{\sbp}{SBProfile}
\newcommand{\mxx}{\ensuremath{M_{xx}}}
\newcommand{\myy}{\ensuremath{M_{yy}}}

\begin{document}

\title{Comparison between \shera\ and \sbp}

\begin{abstract}
This document describes the results of tests of \shera\ and \sbp\ to
ensure that (a) they give consistent results for realistic galaxies
and PSFs and (b) they are applying the shears sufficiently accurately
for the purposes of the GREAT3 challenge.
\end{abstract}

\section{Motivation}

The \shera\ software was designed specifically to simulate 
ground-based data using realistic training data from HST, and was
written in IDL.  Recently \sbp, which was designed as a more
general-purpose image simulator, was extended to include \shera-like
functionality using the SBPixel class to represent general, non-parametric galaxy and
PSF images.  We
want to ensure that the two codes give consistent results, and that
the shears are applied to sufficiently high accuracy for use in GalSim
and GREAT3.

\subsection{The approaches}

There are a few differences between the approaches in \shera\ and in
\sbp\ when mapping a COSMOS galaxy to a sheared one at lower
resolution.  

\begin{itemize}
\item The main difference is that \sbp\ supports any ratio of output
  to input pixel scale, which requires it to interpolate the target
  PSF in Fourier space.  In contrast, \shera\ has more restrictive
  rules about the ratio of output to input pixel scale (requiring that
  $5\times$ (output pixel scale)$/$(input pixel scale) be an integer),
  but this requirement allows the code to avoid explicitly
  interpolating the target PSF.  Instead, it is implicitly
  interpolated, via padding in Fourier space, which is equivalent to
  the ideal sinc interpolation in real space.  Clearly, we require the
  more general \sbp\ treatment for GalSim, and it will be interesting
  to compare these approaches.
\item Due to the aforementioned rules about ratios of output to input
  pixel scale, and method of dealing with target PSFs, \sbp\ requires
  images with linear dimension $2^n$ or $3\times 2^n$ for FFTs, but \shera\ has rather
  more arcane rules that require the galaxy image and the target PSF
  to be related by a factor of output vs. input pixel scale.
\item \shera\ suffers from some annoying centroiding issues
  due to the use of IDL convolution routines.
\item \shera\ and \sbp\ have different flux normalization conventions
  due to their different purposes.  In particular, \sbp\ seeks to
  preserve the surface brightness whereas \shera\ asks the user to specify a target
  flux normalization.
\item \shera\ is carrying out a pseudo-deconvolution whereas \sbp\ is
  carrying out a full deconvolution.  This should not matter as long
  as the pseudo-deconvolution properly looks like a deconvolution for
  all $k$ that are not killed off by the target PSF.
\item The final resampling (via interpolation or subsampling) takes
  place in Fourier domain for \sbp, in real space for \shera.
\end{itemize}

Note that the standard interpolation method that is used by the two is
the same: cubic interpolation (a windowed
appoximation to a sinc based on cubic polynomials, using 16 points for
a 2d image).  But the place where it is applied differs, which could
be important.  However, this difference in interpolation only relates to the
PSF-matching and resampling;
for both pieces of software, the lensing shear is applied via cubic
interpolation in Fourier space.

\subsection{Issues}

Comparison of results between the two codes is complicated by the
different conventions for centroiding and flux normalization, and by
the different rules for array sizes. For our initial tests, to allow for compatible input
image sizes between the codes, we require output vs. input pixel size
$=2^n$.  We chose $n=3$, so the ground-based telescope for which we
carry out the tests has $0.24$ arcsec pixels.  (Note that we still use
SDSS PSFs when we need real ones.  We are simply lying to the code
about the pixel size for those PSFs, for our own convenience.)  Also,
since output vs. input pixel sizes are an integer, the final
resampling to the target pixel size is simply a resampling process.
We will eventually want to test the effects of resampling via
interpolation, but that test can take place at a later stage.

\section{The PSF test and non-sheared galaxy tests}

We begin with a test for which we know the answer: we choose some
typical COSMOS PSF as our ``galaxy,'' so that when we carry out
PSF-matching, the final ``galaxy'' image should equal the target PSF.
This test does not involve shearing, it only tests PSF-matching for
rather arbitrary images.  We also carry out the PSF-matching for three
sample galaxies, where the true answer is not known.  

For this test, we compare three types of outputs: we compare the final
image on the target pixel scale; and we compare it on the original pixel
scale (after PSF-matching, which includes the target pixel response
function).  Since the ratio of pixel scales is an integer, the
first image comes from strict subsampling of the second image.

As an aside, we also compare a real-space version of the target PSF
interpolated to the original COSMOS pixel scale, as a test of our
interpolators. This test 
suggests that our cubic convolution implementations do essentially the
same thing: the moments of the interpolated PSFs agree at the
$10^{-5}$ level.

In the initial version of the 1st and 2nd tests (final simulated image
on target and initial pixel scales), we found that the first test gave
results that agree down to the $3\times 10^{-4}$ level for the sum of
$\mxx+\myy$, and $3 \times 10^{-6}$ in ellipticity -- but the second
test gave differences of order a few percent, with \sbp\ outputs being
larger and rounder than the \shera\ outputs.  

Gary's theory is that
this relates to the fact that his Fourier-domain treatment includes
operations to represent the (by default cubic) interpolation of the
target PSF.  The cubic interpolation effectively broadens the PSF
slightly, so \sbp\ gives broader outputs at the initial pixel scale,
but not at the final pixel scale (where the resampling essentially
means there is no interpolation).  To test the theory that
this difference comes from the \sbp\ inclusion of the cubic
interpolant as part of the definition of the PSF, Gary replaced the
cubic interpolation of the target PSF with an $N=7$ Lanczos filter.
This filter should more closely approximate what \shera\ does, i.e. no
interpolation of the target PSF, but rather padding and embedding it
into some larger array, which is equivalent to sinc interpolation.
When he does this, the $\mxx+\myy$ from \shera\ and \sbp\ at the
initial COSMOS pixel scale agree to a few $\times 10^{-4}$, and the
ellipticities agree to the fifth decimal place.  

Conclusion for GalSim: we need to keep in mind that interpolation of
the target PSF (either implicitly via FTing and padding, or explicitly
in real or Fourier space) can amount to a different definition of the
PSF, and we must decide what definition to use.

\section{Shear}

We have to test the accuracy of applied shears.  There is no serious
reason to expect the methods to disagree with each other (given that
we established that they carry out cubic interpolation nearly
identically), but this test is still important, particularly a test of
absolute rather than relative accuracy.

\subsection{Relative accuracy}

Real galaxies, \sbp\ vs. \shera.

\subsection{Absolute accuracy}

Parametric models, both \sbp\ and \shera\ compared to a known
reference that can be sheared analytically.

\section{Resampling}

Once we understand the results of the above tests, we need to do some tests that 
require resampling to non-integer ratio pixel scales.

\section{Outstanding issues and conclusions}

As we found in the first test, we need to be sure to understand the impact of
resampling with a given interpolant in Fourier space (which gives
errors periodic in $k$, and real-space ghosts) or real space (which
gives errors periodic in $x$, and Fourier-space ghosts), and the
impact on the effective PSF.

A point that Gary raised to consider for the future: are our images
giving surface brightness (the assumption of \sbp) or flux per pixel
(the assumption of \shera, which means to match to some known object
flux in ground-based data)?  And when we add magnification to the
pixel, do we want to conserve flux or surface brightness?  (\sbp\ does
the latter.  We should make this choice clear and decide whether it
makes sense for us for GalSim.  I think it does.)



\end{document}
